---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns

In SQL, you can filter columns by using SELECT and rows by using WHERE. Let's looks at an example using the Lahman baseball database. First, we should load the Lahman package. We will also need the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at an example. Suppose we want to see the home run totals for the 1927 Yankees. We could write:

```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

Here are all of the instances where a Yankee has hit 40 or more home runs.
```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

Here are all of the instances where a player hit more than 40 home runs but had less than 60 strikeouts.
```{r}
query<-"SELECT playerID, teamID, yearID, HR, SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)
```

Here are all of the times that a Phillie in the 1970s hit more than 30 home runs.
```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE teamID='PHI' and yearID>=1970 and yearID<=1979 and HR>30"
sqldf(query)
```

### ORDER BY

With SQL, you can also order the rows in the output. For example, suppose you watned to see every instance of a player hitting more than 50 home run. But you would like the players with the most home runs to be at the top of this list. We could do this in the following way:
```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE HR>50
ORDER BY HR DESC"

sqldf(query)
```

Here is an example of all players with less than 10 strikeouts with at least 400 at-bats.
```{r}
query<-"SELECT playerID, teamID, yearID, SO, AB FROM Batting
WHERE SO<10 and AB>=400
ORDER BY SO"
sqldf(query)
```

###GROUP BY

With SQL, you can group columns and find either the sum, average, max, or min. For example, here is how we would find Babe Ruth's career home run totals.
```{r}
query<-"SELECT playerID, sum(HR) FROM Batting
WHERE playerID='ruthba01'
GROUP BY playerID"
 
sqldf(query)
```

Here is an example of how you would find the career home run totals of any player greater than 600 and ordered in descending order.
```{r}
query<-"SELECT playerID, sum(HR) FROM Batting
GROUP BY playerID
HAVING sum(HR)>600
ORDER BY sum(HR) DESC"

sqldf(query)
```

Here is an example of how you would find the average home run totals for a season with players that average over 30.
```{r}
sqldf(query)
query<-"SELECT playerID, avg(HR) FROM Batting
GROUP BY playerID
HAVING avg(HR)>30
ORDER BY avg(HR) DESC"
sqldf(query)
```

