---
title: "SQL"
output: html_document
---

## Filtering Rows and Columns

In SQL, you can filter columns by using SELECT and rows by using WHERE. Let's looks at an example using the Lahman baseball database. First, we should load the Lahman package. We will also need the sqldf package.

```{r message=FALSE, warning=FALSE}
library(Lahman)
library(sqldf)
```

Let's take a look at an example. Suppose we want to see the home run totals for the 1927 Yankees. We could write:

```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE teamID='NYA' and yearID=1927"
sqldf(query)
```

Here are all of the instances where a Yankee has hit 40 or more home runs.
```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE teamID='NYA' and HR>=40"
sqldf(query)
```

Here are all of the instances where a player hit more than 40 home runs but had less than 60 strikeouts.
```{r}
query<-"SELECT playerID, teamID, yearID, HR, SO FROM Batting
WHERE HR>40 and SO<60"
sqldf(query)
```

Here are all of the times that a Phillie in the 1970s hit more than 30 home runs.
```{r}
query<-"SELECT playerID, teamID, yearID, HR FROM Batting
WHERE teamID='PHI' and yearID>=1970 and yearID<=1979 and HR>30"
sqldf(query)
```
